Enriching Object Detection with 2D-3D Registration\\and Continuous Viewpoint Estimation

### 2015-03-30 approach.tex
Major approach modification don. minor mcmc and overview
Approach major whiten to LDA update
MCMC minor math update
experiment, major PASCAL12 table to scriptsize

We used an off-the-shelf rendering engine that can create a realistic rendering
and a depth map. The CAD models we used contain color, texture and material
information such as transparency and reflectance to make the rendering as
realistic as possible to simulate the natural image statistics. An rendering
example that we used in our experiments is in Fig. \ref{fig:rendering}. Along
with a rendering, we extracted a depth map which can later be used for
reconstruction.

To handle intraclass variability, we used various CAD models (sedan, truck,
racing car, hatchback etc) to cover intraclass variability and made renderings
of these CAD models from various viewpoints. The CAD models were cached so that
when we fine-tune 2D-3D matching, we can render the model instantaneously. Our
renderer can continuously vary yaw, pitch, roll and the focal length as well so
that the final fine tuning stage produces accurate the continuous viewpoint, a
focal length and a bounding box.

@@

\subsection{Whitened Histograms of Orientations (WHO)}
\label{sec:who}
Our technique for rendering and generating exemplar template detectors
on-the-fly is drawing from recent work. Hariharan \etal. introduced Whitened
Histograms of Orientations (WHO), which uses feature statistics from natural
images to create a large number of classifiers analytically using Linear
Discriminant Analysis(LDA) rather than training SVM
classifiers\cite{Hariharan12}. The confidence score for data $x_i$ can be defined as

\begin{align}
    S(x_i) = w_{x_t}^T x_i
\end{align}

where $w_{x_t} = \Sigma^{-1} (x_t - \mu)$ is a LDA classifier for a template $x_t$.

@@

Since collecting second order statistics to generate a covariance matrix for
every template shape is expensive, \cite{Hariharan12} proposed easy way to
synthesize covariance $\Sigma$. The method \cite{Hariharan12} assumes
Wide-Sense Stationarity(WSS) of HOG features generated from natural images.
Assumption they made was that the mean $\mu$ and autocovariance $\Gamma$ of HOG
features do not vary with respect to the location. In this paper, we followed
their method to generate $\Sigma$.

@@

\subsection{Whitening Synthesized Templates and Non-Zero WHO}
\label{sec:nzwho}
Our first innovation is `Non-Zero' whitening. When synthesizing detection
templates from rendered images, a common problem is how to handle the
background. If the model is rendered over a natural image background, gradients
in the background will be incorporated into the discriminative template,
potentially causing false-positives if background elements exist in the query
image.

Alternatively, if the background is left textureless (see
Fig.~\ref{fig:rendering}), whitening the resulting HOG template
will produce undesirable background artifacts. These are created when centering
the template (by subtracting the mean $\mu$), where strong negative weights are
introduced in the textureless region (as seen in Fig.~\ref{fig:whocomparison}).
This could result in positive matches being suppressed due to spurious
background gradients.

NZ-WHO removes these artifacts so that the background has no effect on the
template response. Let a vectorized HOG features of a rendering image $x = [ \begin{array}{ccccc}x_1^T & x_2^T & \cdots & x_n^T\end{array} ]^T \in
\mathcal{R}^{nd}$ where $x_i$ is the $i$th HOG cell feature and $n$ is the number HOG cells and $d$ is
the dimension of HOG feature. We create a new vector $x'$ which contains only
the non-zero HOG cell features of $x$. To be specific, let $I_d\in \mathcal{R}^{d\times
    d}$ be the identity matrix, $n'$ be the number of non-zero HOG cells, and a
matrix $S\in \mathcal{R}^{n'd \times nd}$ be the masking matrix that selects
non-zero HOG cells. For instance, a template has  $n=3, n'=2$, and only the
second HOG cell has $0$ norm, then

@@

The matrix selects HOG features that corresponds to the first cell and the third cell. Using the selector $S$, we can define new vectorized HOG features $x' = S x, \mu' = S\mu, \Sigma' = S \Sigma S^T$.
After solving the resulting system (which is now smaller than
in the WHO approach) we find

\begin{equation}
w'=\Sigma'^{-1}(x' - \mu') \label{eq:nz-who},
\end{equation}

which we restore zero cells and reshape the vector into the final NZ-WHO
template that has the same dimension as the original HOG template.

@@

\subsection{Fast Whitening using Conjugate Gradient}
\label{sec:fastwhiten}
Our second innovation lies in enhancing the computational efficiency
of the whitening procedure, making it applicable to an on-the-fly
rendering setting.

It originates from the insight that an iterative
Conjugate Gradient method can be used to whiten a HOG template, which
is much faster than the original procedure proposed
in~\cite{Hariharan12} based on decomposition.

Whitening the non-zero
feature vector $x'$ means a solving the system of linear equations, $\Sigma' w' = (x' -
\mu')$. In
\cite{Hariharan12}, the authors make use of the fact that covariance matrices
are symmetric and positive semidefinite to solve the system via the Cholesky
decomposition with Gaussian Elimination, which requires $O(n^3)$ time.

@@

The covariance matrix for HOG templates is typically
ill-conditioned\cite{Hariharan12}, but we
can add a regularization constant to the diagonal to reduce its condition
number to the point where the conjugate gradient method will converge quickly.
We use a constant of $0.15$, which reduces the condition number from $10^{20}$
to $50$, much smaller than the dimension of the matrix (7000).

As a result, a GPU implementation of conjugate gradient converges in 80
ms when using 250 HOG cells, two orders of magnitude faster than the using
Cholesky factorization with Gaussian Elimination.

We report the real time analysis of whitening using decomposition and
conjugate gradient methods in Fig.~\ref{fig:whotime}. (a) compares the
absolute runtime of the different methods while (b) gives the obtained
speedup. We see that % We compare the speed of each of template generation methods in
% Fig.~\cite{fig:covariancetime_crop} Using naive Cholesky decomposition,
WHO template takes several seconds for realistic template sizes of
several hundred cells, while using the iterative Conjugate Gradient
method, it only takes 80ms. If we use NZ-WHO, we can gain extra speed
up since we only whitens non-zero cells.

@@

\subsection{High Resolution Templates and FFT based Convolution}
\label{sec:fft} 
We generate high resolution templates with more than 250 HOG cells to capture
details of an object to give accurate 2D-3D matching. These large templates
cause computational burden when computing convolution. Though good for
detecting accurate model and pose, convolution of high-resolution templates are
much slower since computation time scales linearly with the number of HOG
cells in the template. To overcome and speed up the convolution, we used FFT
based GPU convolution \cite{Podlozhnyuk} which scales. Briefly, for length $n$ signal and
length $m$ filter, naive convolution takes $O(nm)$ time whereas FFT-based
convolution takes $O\left( (n + m)\log (n+m) \right)$ time. For large
$m$ (high resolution templates), we can gain computational advantage.

>>>>

We used an off-the-shelf rendering engine to generate a realistic rendering
and a depth map. The CAD models we used contain texture and material
information. These make the rendering more realistic and allows us to transfer
natural image statistics to rendered images. Along
with a rendering, we extracted a depth map which can be used for
reconstruction\ref{fig:rendering}.

To handle intraclass and viewpoint variability, we used various CAD models and
made renderings of these CAD models from different viewpoints.  Note that we
continuously vary yaw, pitch, roll and the focal length as well so that the
final fine tuning stage can produce accurate viewpoint estimations
(Sect.~\ref{sec:fine}).

@@

\subsection{Whitened Histograms of Orientations (WHO)}
\label{sec:who}
Our technique for rendering and generating exemplar template detectors
on-the-fly is drawing from recent work. Hariharan \etal. introduced Whitened
Histograms of Orientations (WHO), which uses feature statistics from natural
images to create a large number of classifiers analytically using Linear
Discriminant Analysis(LDA) rather than training SVM
classifiers\cite{Hariharan12}. The confidence score for data $x_i$ can be
defined as $ S(x_i) = w_{x_t}^T x_i$ where $w_{x_t} = \Sigma^{-1} (x_t - \mu)$
is an LDA classifier for a template $x_t$. Since collecting covariance matrices
for all possible template shape is intractable, \cite{Hariharan12} assumed
Wide-Sense Stationarity(WSS) of HOG features and generated a covariance
$\Sigma$ from autocovariance $\Gamma$ collected on a large collection of
natural images. i.e. for a 31 dim HOG feature $x_t$ at location $t$

\begin{align}
    E[x_t] & = \mu\\
    E[(x_{t} - \mu ) (x_{\tau} - \mu)] & = E[(x_0 - \mu)(x_{\tau - t} - \mu)]
\end{align}

In this paper, we followed their method to generate $\Sigma$.

@@

@@

\subsection{Whitening Synthesized Templates and Non-Zero WHO}
\label{sec:nzwho}
Our first improvement is `Non-Zero' whitening. When synthesizing detection
templates from rendered images, a common problem is how to handle the
background. If the model is rendered over a natural image background, gradients
in the background will be incorporated into the discriminative template.

Alternatively, if the background is left textureless (see
Fig.~\ref{fig:rendering}), whitening the resulting HOG template
, strong negative weights are introduced in the textureless region (by
subtracting the mean $\mu$) as seen in Fig.~\ref{fig:whocomparison}.
This could result in positive matches being suppressed due to spurious
background gradients.

NZ-WHO removes these artifacts so that the background has no effect on the
template response. Let a vectorized HOG features of a rendering image $x = [
\begin{array}{ccccc}x_1^T & x_2^T & \cdots & x_n^T\end{array} ]^T \in
\mathcal{R}^{nd}$ where $x_i$ is the $i$th HOG cell feature and $n$ is the
number HOG cells and $d$ is the dimension of HOG feature. We create a new
vector $\bar{x}$ which contains only the non-zero HOG cell features of $x$. To be
specific, let $I_d\in \mathcal{R}^{d\times d}$ be the identity matrix, $\bar{n}$ be
the number of non-zero HOG cells, and a matrix $S\in \mathcal{R}^{\bar{n}d \times
    nd}$ be the masking matrix that selects non-zero HOG cells. For instance, a
template has  $n=3, \bar{n}=2$, and only the second HOG cell has $0$ norm, then

@@

The matrix selects HOG features that corresponds to the first cell and the third cell.

Using the selector $S$, we can define new vectorized HOG features $\bar{x} = S x,
\bar{\mu} = S\mu, \bar{\Sigma} = S \Sigma S^T$. After solving the resulting system (which
is now smaller than in the WHO approach) we find

\begin{equation}
    \bar{w}=\bar{\Sigma}^{-1}(\bar{x} - \bar{\mu}) \label{eq:nz-who},
\end{equation}

To speed up the convolution, we restore zero cells and reshape the vector $\bar{w}$
and compute convolution.

@@

\subsection{Fast Whitening using Conjugate Gradient}
\label{sec:fastwhiten}
% To speed up LDA synthesize, we introduced Conjugate Gradient method.
% for solving the matrix inversion Eq.~\ref{eq:nz-who}.
% This process speeds up , making it applicable to an on-the-fly rendering setting.
% It originates from the insight that an iterative Conjugate Gradient method can be used to whiten a HOG template, which is much faster than the original procedure proposed in~\cite{Hariharan12} based on decomposition.

Synthesizing an LDA template Eq.~\ref{eq:nz-who} requires solving the system of
linear equations, $\bar{\Sigma} \bar{w} = (\bar{x} - \bar{\mu})$. In
\cite{Hariharan12}, the authors make use of the fact that covariance matrices
are symmetric and positive semidefinite to solve the system via the Cholesky
decomposition with Gaussian Elimination, which requires $O(n^3)$ time.

@@

The covariance matrix for HOG templates is typically
ill-conditioned\cite{Hariharan12}, but adding a small regularization constant to
the diagonal reduces its condition number.
We use a constant of $0.15$, which reduces the condition number from $10^{20}$
to $50$, much smaller than the dimension of the matrix (7000).

As a result, a GPU implementation of conjugate gradient converges in 60
ms when using 250 HOG cells on Nvidia GTX660, two orders of magnitude faster
than the using Cholesky factorization with Gaussian Elimination.

We report the real time analysis of whitening using decomposition and
conjugate gradient methods in Fig.~\ref{fig:whotime}. (a) compares the
absolute runtime of the different methods while (b) gives the obtained
speedup. We see that % We compare the speed of each of template generation methods in
% Fig.~\cite{fig:covariancetime_crop} Using naive Cholesky decomposition,
WHO template takes several seconds for realistic template sizes of
several hundred cells, while using the iterative Conjugate Gradient
method, it only takes 60ms. If we use NZ-WHO, we can gain extra speed
up since we only whitens non-zero cells.

@@

\subsection{High Resolution Templates and FFT based Convolution}
\label{sec:fft}

We generate high resolution templates with more than 250 HOG cells to capture
details of an object to give accurate 2D-3D matching. These large templates
cause computational burden when computing convolution. Though good for
detecting accurate model and pose, convolution of high-resolution templates are
much slower since computation time scales linearly with the number of HOG
cells in the template. To overcome and speed up the convolution, we used FFT
based GPU convolution \cite{Podlozhnyuk} which scales. Briefly, for length $n$
signal and length $m$ filter, naive convolution takes $O(nm)$ time whereas
FFT-based convolution takes $O\left( (n + m)\log (n+m) \right)$ time. For large
$m$ (high resolution templates), we can gain computational advantage.

### 2015-03-30 mcmc.tex
compressed
minor changes
Major approach modification don. minor mcmc and overview
Approach major whiten to LDA update
MCMC minor math update
experiment, major PASCAL12 table to scriptsize

This algorithm changes the current state $p$ to a new state $q$ based
on the acceptance probability
\begin{align}
    A(p \rightarrow q) & =  min\left( 1,  \frac{P(q) g(q \rightarrow p)}{P(p) g(p \rightarrow q)}\right).
\end{align}
We define $3$ different types of moves that can alter the state
(Fig.~\ref{fig:moves}), {\em (i)} changing the focal length $f$, {\em
(ii)} one of the rotational pose parameters $v_i$, and {\em (iii)} CAD model
index $m$. For (i) and (ii), we use Gaussian proposal distributions, and a uniform
distribution over model changes for (iii). For scale and translation, we
implicitly compute all possible translation and scale by convolving $w(q)$ to
HOG pyramid.
\begin{align}
    g(p \rightarrow p + v_i') & \sim \mathcal{N}(\mu = p_{v_i},\sigma = \sigma_v) \quad i \in \{1,2,3\}\\
    g(p \rightarrow p + f') & \sim \mathcal{N}(\mu = p_{f},\sigma = \sigma_f)\\
    g(p \rightarrow p + m') & \sim c \delta(p_m) + (1-c) Unif(1,M),
\end{align}
where $g$ is the proposal distribution. % and $P$ is the distribution of which we
%defined using Gaussian distribution around previous state. $A$ is the
%acceptance probability where we accept new proposal with $A(x \rightarrow x')$.
%
% One can also think of the image as an image cropped by detection
% bounding box. We synthesize covariance matrix and create NZ-WHO
% template $w(x)$ for particular viewpoint, focal length and CAD model.
% In sum, for new proposal $x'$, we make a NZ-WHO template on the fly 
% We found out that it requires good initialization to converge to
% local minima.

@@

% there are several advantages that was not possible such as
% \textbf{on-the-fly} template generation and making large number of
% templates
\begin{figure}[t]
\centering
    \includegraphics[width=0.7\linewidth]{tuning2} \\ [-5pt]
    \caption{Metropolis Hastings move types (Section~\ref{sec:fine}).}
 \label{fig:moves}
\end{figure}

\paragraph{Initialization.}
Since the Metropolis-Hastings algorithm is susceptible to getting
stuck in local optima, we need a good initialization in order to
increase the probability to find a solution that is close to the
global optima. We achieve this by first running a discrete,
pre-trained version of our algorithm (i.e., an ensemble of NZ-WHO
templates) in order to get promising candidate 2D bounding boxes and
poses to start from. We then initialize $x_0$ for each of these candidates.

>>>>

This algorithm changes the current state $p$ to a new state $q$ based
on the acceptance probability
\begin{align}
    A(p \rightarrow q) & =  min\left( 1,  \frac{P(q) g(q \rightarrow p)}{P(p) g(p \rightarrow q)}\right).
\end{align}
We define $3$ different types of moves that can alter the state
(Fig.~\ref{fig:moves}), {\em (i)} changing the focal length $f$, {\em
(ii)} one of the rotational pose parameters $v_i$, and {\em (iii)} CAD model
index $m$. For (i) and (ii), we use Gaussian proposal distributions, and a uniform
distribution over model changes for (iii). For scale and translation, we
implicitly compute all possible translation and scale by convolving $w(q)$ to
HOG pyramid.
\begin{align}
    g(p \rightarrow p + v_i') & \sim \mathcal{N}(\mu = p_{v_i},\sigma = \sigma_v) \quad i \in \{1,2,3\}\\
    g(p \rightarrow p + f') & \sim \mathcal{N}(\mu = p_{f},\sigma = \sigma_f)\\
    g(p \rightarrow p + m') & \sim (1-c) \delta(p_m) + c Unif(1,M),
\end{align}
where $g$ is the proposal distribution. % and $P$ is the distribution of which we
%defined using Gaussian distribution around previous state. $A$ is the
%acceptance probability where we accept new proposal with $A(x \rightarrow x')$.
%
% One can also think of the image as an image cropped by detection
% bounding box. We synthesize covariance matrix and create NZ-WHO
% template $w(x)$ for particular viewpoint, focal length and CAD model.
% In sum, for new proposal $x'$, we make a NZ-WHO template on the fly 
% We found out that it requires good initialization to converge to
% local minima.

@@

% there are several advantages that was not possible such as
% \textbf{on-the-fly} template generation and making large number of
% templates
\begin{figure}[t]
\centering
    \includegraphics[width=0.7\linewidth]{tuning2} \\ [-5pt]
    \caption{Search space that our fine-tuning stage handles (Section~\ref{sec:fine}).}
 \label{fig:moves}
\end{figure}

\paragraph{Initialization.}
Since the objective function is non-convex, we need a good initialization in order to
increase the probability to find a solution that is close to the
global optima. We achieve this by first running a discrete,
pre-trained version of our algorithm (i.e., an ensemble of NZ-WHO
templates) in order to get promising candidate 2D bounding boxes and
poses to start from. We then initialize $x_0$ for each of these candidates.
